---
title: "Instructor guide to using {sandpaper}"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Instructor guide to using {sandpaper}}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(sandpaper)
```

## Introduction

This guide provides an overview of how to use the {sandpaper} package to create and maintain
lessons, mainly targeted at instructors.
**For a complete guide to using {sandpaper}, head over to
<https://carpentries.github.io/sandpaper-docs/>.**

## Creating a new lesson

### Setting up the directory structure

To create a lesson with {sandpaper}, use the `create_lesson()` function:

```r
sandpaper::create_lesson("~/Desktop/r-intermediate-penguins")
```

This will create folder on your desktop called `r-intermediate-penguins` with
the following structure:

```
|-- .gitignore               # - Ignore everything in the site/ folder
|-- .github/                 # - Scripts used for continuous integration
|   `-- workflows/           #
|       |-- deploy-site.yaml # -   Build the source files on github pages
|       |-- build-md.yaml    # -   Build the markdown files on github pages
|       `-- cron.yaml        # -   reset package cache and test
|-- episodes/                # - PUT YOUR MARKDOWN FILES IN THIS FOLDER
|   |-- data/                # -   Data for your lesson goes here
|   |-- figures/             # -   All static figures and diagrams are here
|   |-- files/               # -   Additional files (e.g. handouts)
|   `-- introduction.Rmd     # -   Lessons start with a two-digit number
|-- instructors/             # - Information for Instructors
|-- learners/                # - Information for Learners
|   `-- setup.md             # -   setup instructions (REQUIRED)
|-- profiles/                # - Learner and/or Instructor Profiles
|-- site/                    # - This folder is where the rendered markdown files and static site will live
|   `-- README.md            # -   placeholder
|-- config.yaml              # - Use this to configure commonly used variables
|-- CONTRIBUTING.md          # - Carpentries Rules for Contributions (REQUIRED)
|-- CODE_OF_CONDUCT.md       # - Carpentries Code of Conduct (REQUIRED)
|-- LICENSE.md               # - Carpentries Licenses (REQUIRED)
`-- README.md                # - Introduces folks how to use this lesson and where they can find more information.
```

Once you have your site set up, you can add your RMarkdown files in the episodes
folder. By default, they will be built in alphabetical order, but you can use
the `set_episodes()` command to build the schedule in your `config.yaml` file:

```r
s <- sandpaper::get_episodes()
sandpaper::set_episodes(order = s, write = TRUE)
```

### Adding new episodes

You can add new episodes, using

```r
sandpaper::create_episode("plotting")
```

This will add a new `episodes/plotting.Rmd` file to your lesson with some example content. You can
then edit this file to add your own content.

### Episode structure

See [Introduction to The Carpentries Workbench: Episode Structure](https://carpentries.github.io/sandpaper-docs/episodes.html).

### Previewing your new lesson

After you created your lesson, you will want to preview it locally. First, make sure that you are in
your newly-created repository and then use the following command:

```r
sandpaper::serve()
```

This function will open a preview window in RStudio or your browser and will **update
automatically** as you work on the lesson.

If you are using RStudio, you can use the following keyboard shortcuts:

- Render and preview the whole lesson: `ctrl/cmd + shift + B`
- Render and preview an episode: `ctrl/cmd + shift + K`

### Pushing to GitHub

Once you are happy with your lesson, you can push it to GitHub to run the automated lesson building
workflows and to publish your lesson on GitHub pages. You can set up a GitHub repository and activate GitHub pages using the [{usethis}](https://usethis.r-lib.org/) package:

```r
usethis::use_github(organisation = "<ORG-NAME>", private = FALSE)
usethis::use_github_pages()
```

Replace `"<ORG-NAME>"` with the name of your GitHub organisation or set it to `NULL` to use your
personal account. You can also make the repository private by setting `private = TRUE`. This will
create a new GitHub repository at `https:://github.com/<ORG-NAME>/r-intermediate-penguins` and will
set up GitHub pages at `https://<ORG-NAME>.github.io/r-intermediate-penguins`. It will take a couple
of minutes for the GitHub Action workflows to validate, build and deploy the lesson. You can track
the progress at `https:://github.com/<ORG-NAME>/r-intermediate-penguins/actions`.

## Modifying an existing lesson

To work on an existing lesson already hosted on GitHub,
[fork and clone the repository](https://docs.github.com/en/get-started/quickstart/fork-a-repo) to
your computer using your method of choice. When working in R, you can do this easily with the [{usethis}](https://usethis.r-lib.org/) package:

```r
## Using an example repo from the Carpentries
usethis::create_from_github("carpentries/sandpaper-docs", "~/Documents/Lessons/")
```

After editing the content and previewing with `sandpaper::serve()`, you should first **save and
commit your changes** with `git`. When working in RStudio, you can do this in the Git tab in the top
right corner of the screen:

1. Click the "Git" tab in the top right corner of the screen
2. Check the "Staged" box for the file(s) you want to commit
3. Type a concise but descriptive commit message in the "Commit message" box
4. Click "commit"

After committing your changes, you can push them to GitHub using the "Push" button in the Git tab.
Note that it's often
[a good idea to to make several, small commits](https://happygitwithr.com/repeated-amend.html#rock-climbing-analogy)
rather than one large commit. You also don't need to push after each commit.

For a comprehensive guide on using Git and GitHub with R, see [Happy Git and GitHub for the useR](https://happygitwithr.com/).

## Managing and updating the package cache

>*Note: this is only relevant for lessons with __generated content__. I.e. lessons based on RMarkdown episodes*

Full guide: [Building Lessons With A Package Cache • sandpaper](https://carpentries.github.io/sandpaper/articles/building-with-renv.html).

{sandpaper} can set up a package cache for your lesson to create a reliable setup that ensures the
same package versions are used to build your lesson, both locally and on the GitHub Actions runners.

To make use of the package cache, you need to *explicitly* give {sandpaper} permission to create and use a cache with:

```r
sandpaper::use_package_cache()
```

From this point on, {sandpaper} will detect any **R packages** used throughout the lesson with
`library(package)` or `package::some_function()` and add them to the cache.

>*Note: this currently does not work for __Python packages__, which still need to be manually installed for the lesson with `sandpaper::py_install()`.*

Now every time you build the lesson (either locally or on GitHub Actions), {sandpaper} will check
that all the packages used in the lesson are available in the cache. If a package is missing,
{sandpaper} will install it and add it to the cache.
